---
title: "Homework 5--DS 700"
author: "Abra Brisbin"
date: "12/17/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1:  R as a scientific calculator
### 1a.
Use the `c` function to make a vector containing the numbers -1, 0, 2, 5, 17.  Store your vector in the variable x.
```{r}
x <- c(-1,0,2,5,17)
```

Use square brackets [ ] to tell R to display the 3rd element of x.
```{r}
x[3]
```
Display the 6th element of x.  What happens?
```{r}
x[6]
```

Try running the following code.
```{r}
x[-1]
x[-2]
x[-3]
```
What does the negative sign inside square brackets do?
<!-- It seems to remove the specific element at that point in the vector. -1 removes the first element of a vector, -2 the second element of a vector, etc... -->
### 1b.
*Without using `c`*, make a vector of numbers from -3 to 13, in increments of 4.  Store your vector in the variable y, and display the contents of y.

```{r}
y <- seq(-3,13, by=4)
```

Add 5 to each element of y and store the results in the variable z.  Display your results.  
```{r}
z=y+5
```

Take the square root of each element of y.  (Hint:  You can Google "square root r" to find the function to take the square root.)
```{r}
sqrt(y)
```
Add each element of x to the corresponding element of y.  Store your results back in the variable y.  (This will overwrite the old contents of y.)  Display your results.
```{r}
y=x+y
```

### 1c.

Run the following code to create a vector in a new way:
```{r}
my_subset = 2:4
my_subset
```

What does the `:` do?  What are two other ways we could create the same vector?

Use square brackets to extract elements 2, 3, and 4 of `x`.  Your code in the next code cell should not include any numbers.
```{r}
x[2:4]
```
Use square brackets to extract elements 1 and 5 of `x`.   Your code in the next code cell should not include any numbers.
```{r}
x[c(1,5)]
```


## Problem 2:  Analyzing a data set
This problem will guide you through reading a data set into R and understanding it.

### 2a.
Install the following packages, if you do not already have them installed:

- ggformula
- dplyr
- readr

Remember to do this in the console (lower-left panel of RStudio) or the Packages tab (in the lower-right panel), not in your .Rmd document.  Trying to install packages from the .Rmd document can result in an error about "trying to use CRAN without a mirror".

After installing the packages, use the `library` function to load them in the code cell below.  (Normally, I like to load all my packages at the start of an .Rmd file.  That way, if I need to run the .Rmd file on a new computer, it's easy to see what I need to install.)

```{r}
library(ggformula)
library(dplyr)
library(readr)
```

Which objects are masked by objects in `dplyr`?  Should we be concerned about these?
<!-- intersect, setdiff, setequal, union -->
<!-- filter, lag -->

### 2b.
Read Wisconsin_income.csv into R.  Call the data frame `wi`.
```{r}
wi <- read_csv("~//Documents/R/Wisconsin_income.csv")
head(wi)
class(wi$AGEP)
class(wi$COW)
class(wi$JWMNP)
class(wi$LANX)
class(wi$MAR)
class(wi$SEX)
class(wi$WKHP)
class(wi$DIS)
class(wi$PERNP)
class(wi$Hispanic)
class(wi$Stem)
class(wi$Education)
class(wi$Region)
class(wi$CIT2)
class(wi$RAC)
```
Which variables have been read in as categorical?
Region and Stem

Use a text editor (such as Notepad++) to open the data dictionary.  Which variables should be categorical, but have been read in as numeric?
CIT2,DIS,RAC,Hispanic,SEX,COW,LANX,MAR

Tell R to treat those variables as categorical.
```{r}
wi$CIT2<-as.factor(wi$CIT2)
wi$SEX<-as.factor(wi$SEX)
wi$DIS<-as.factor(wi$DIS)
wi$RAC<-as.factor(wi$RAC)
wi$Hispanic<-as.factor(wi$Hispanic)
wi$COW<-as.factor(wi$COW)
wi$LANX<-as.factor(wi$LANX)
wi$MAR<-as.factor(wi$MAR)
```

### 2c.
Sometimes you'll need to create your own code cells.  Start a code cell below by typing 3 backticks (on the key to the left of the number 1) and then {r}.  End the cell on a new line by typing 3 backticks again.
```{r}
ncol(wi)
nrow(wi)
```


Inside the cell you created, write a line of code to display the number of rows and columns of the data.

In the code cell below, write code to display the first 6 rows of data.
```{r}
head(wi)
```

Use the first 6 lines of data and the data dictionary to help you answer the question:  In this data set, what does 1 observational unit represent?

### 2d.
Make a boxplot of earnings, grouped by class of worker.
```{r}
boxplot(PERNP ~ COW, data = wi)
```

When most observations of a variable are clustered around a low value, and most of the outliers are on the high end of the distribution, we say the variable is "right-skewed".  When a variable is right-skewed, it can sometimes be helpful to reduce the skew by taking the log.  

Run the following code cell to make a boxplot of log(earnings), grouped by class of worker.

```{r}
gf_boxplot(log(PERNP) ~ COW, data = wi)
```

Which class of worker has the lowest-paid person in the data set?  

Which class of worker has the lowest median income?  (Note that in a boxplot, the median is denoted by the thick line in the middle of the box.)

## 2e.
Make a scatterplot of earnings as a function of hours worked per week.  
```{r}
attach(wi)
plot(WKHP, PERNP, main="Relationship Between Earnings and Work",
   xlab="Hours Worked Per Week", ylab="Total Person's Earnings", pch=19)

abline(lm(PERNP~WKHP), col="blue") # regression line (y~x)
lines(lowess(WKHP,PERNP), col="green") # lowess line (x,y)
```

Comment on the relationship between the variables.
<!-- It appears that the fit line is positive but curved at a certain point that hours don't matter as much to total earnings. -->

Let's see if we can get a more informative graph by reducing the skew in earnings.  Make a scatterplot of the relationship between log(earnings) and hours worked per week.
```{r}
attach(wi)
plot(WKHP, log(PERNP), main="Relationship Between Earnings and Work",
   xlab="Hours Worked Per Week", ylab="Total Person's Earnings", pch=19)

abline(lm(log(PERNP)~WKHP), col="blue") # regression line (y~x)
lines(lowess(WKHP,log(PERNP)), col="green") # lowess line (x,y)
```

Describe the relationship between the variables.
<!-- It is a positive relations -->
Add a line of best fit to the graph.  Is its slope positive or negative?  What does this tell you, in the context of this data set?
```{r}

```

