---
title: "DS 700 Lesson 6 Homework"
author: "Abra Brisbin"
date: "11/22/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Factors affecting the price of used cars

Load necessary packages here.
```{r}
library(readr)
library(ggplot2)
library(ggformula)
library(dplyr)
```


## Problem 1:  Graphical analysis
###1a.	Download Cars 2005.csv and load the data into R.  Display the first few lines of data.

(Dataset:  "Car Data," submitted by Shonda Kuiper, Grinnell College. Dataset obtained from the Journal of Statistics Education (http://www.amstat.org/publications/jse). Accessed 3 June 2015. Used by permission of author.)


```{r}
setwd('C:/Users/aeabr9256/Documents/R')
Cars_2005 <- read_csv("Cars_2005.csv")
head(Cars_2005)
```

***Note:**  All of parts b-k should be done using functions from `ggformula`.

###1b. The variables Cruise, Sound, and Leather use 1 to represent "yes" (the car has cruise control, an upgraded sound system, or a leather interior) and 0 to represent "no".  Therefore, these variables are categorical--but R treats them as numeric, because they look like 0's and 1's.
The function `as.factor()` will return a categorical variable version of its argument (the variable you put inside the parentheses).  Use `mutate` and `as.factor` to convert the variables Cruise, Sound, and Leather into factor variables.  (It may help to review the Reading data into R video in Lesson 5.)
```{r}
Cars_2005 <- Cars_2005 %>%
  mutate(Cruise = as.factor(Cruise),
         Sound = as.factor(Sound),
         Leather = as.factor(Leather),
         Liter = as.factor(Liter),
         Cylinder = as.factor(Cylinder))
```

###1c. Make a graph displaying the relationship between Price and Liter (a measure of engine size).  
```{r}
gf_boxplot(Price ~ Liter, data = Cars_2005)

gf_point(Price ~ Liter, data = Cars_2005, shape = ~ Leather, color = ~ Leather) %>% 
  gf_lm() %>% 
  gf_labs(title = "How Engine size impacts Price",
          subtitle = "Leather or No Leather",
          x = "Liter",
          y = "Price")
```

- **Write a sentence** explaining which variable you chose as `x`, the independent variable, and why.
I chose liter as the independent variable (x) because the size of the engine has an impact on the dependent variable (y) of price

- **Write a sentence** describing the relationship between these two variables.
It appears to be a positive relationship between the two variables

###1d. Modify your graph to include the variable Leather, in addition to Price and Liter. **Include** an informative title for your graph.    There should be one set of axes.
```{r}
gf_boxplot(Price ~ Liter| Leather, data = Cars_2005) %>%
  gf_labs(title = "How Engine size impacts Price",
          subtitle = "Leather or No Leather",
          x = "Liter",
          y = "Price")

```
- **Write a sentence** explaining which attribute you mapped to Leather, and why.

###1e. Make a graph showing the distribution of Price.  Use vertically-stacked facets to display the price for cars with and without leather interiors.  There should be one set of axes.
```{r}
gf_histogram(~ Price, data = Cars_2005) %>%
  gf_facet_grid(Leather ~ .)
```

###1f. Make a graph showing the distribution of Price, with side-by-side bars for cars with and without leather.   There should be one set of axes.
```{r}
gf_histogram(~ Price, fill =~ Leather, position = position_dodge(), data = Cars_2005)

```

###1g. Make another type of graph showing the relationship between Price and Leather.
```{r}
gf_boxplot(Price ~ Leather, data = Cars_2005)
```

- **Write a sentence** describing the relationship between Price and Leather.
In this graph, more cars seems to have leather than do not and as you get into more expensive cars they have leather as a standard.

- In your opinion, which of the graphs in parts e-g is the most effective for communicating the relationship between Price and Leather?
I believe the side by side histogram gave me the best understanding of the data set in comparison to the rest.


###1h. Make *one* graph showing the relationship between Type (Sedan, Convertible, etc.) and Leather.
```{r}
gf_bar(~ Type, fill =~ Leather, position = position_dodge(), data = Cars_2005)
```

- **Write 1-3 sentences** describing the relationship between Type and Leather.
There are way more cars with leather than without no matter the type of car.

## Problem 2:  Regression analysis
### 2a.  Use linear regression to model Price as a function of Liter.

```{r}
price.liter <- lm(Price ~ Liter, data = Cars_2005)
summary(price.liter)

gf_point(Price ~ Liter, data = Cars_2005) %>%
  gf_smooth(Price ~ Liter, method = "lm", col = "blue", data = Cars_2005)

```

What is the equation of the line of best fit?

y= 4990.4x+6185.8

```{r}
Cars_2005 <- Cars_2005 %>%
  mutate(Liter = as.numeric(as.character(Liter)))

gf_point(Price ~ Liter, data = Cars_2005) %>%
  gf_smooth(Price ~ Liter, method = "lm", col = "blue", data = Cars_2005)

price.liter <- lm(Price ~ Liter, data = Cars_2005)
summary(price.liter)
         
```

**Write a sentence** interpreting the slope of the line of best fit.

As 1 liter is added to the engine size of a vehicle, the price increases by $4990

### 2b. Graph the residual plots of the linear model.
```{r}
par(mfrow = c(2,2))
plot(price.liter)

```
In 2-4 sentences, comment on the appropriateness of the linear model. 

### 2c. Use linear regression to model log(Price) as a function of Liter.

```{r}
Cars_2005 <- Cars_2005 %>%
  mutate(logPrice=log(Price))

logmodel = lm(logPrice ~ Liter, data = Cars_2005)


```

Write an equation for estimated Price (not log(Price)).

### 2d. Graph the residual plots of the linear model for log(Price).  In 2-4 sentences, comment on the appropriateness of the model. 
```{r}
plot(logmodel)
logmodel
```

### 2e. Make a scatterplot of log(Price) as a function of Liter.  Plot the line of best fit on the graph.
```{r}
gf_point(logPrice ~ Liter, data = Cars_2005) %>%
  gf_smooth(logPrice ~ Liter, method = "lm", col = "blue", data = Cars_2005)

logprice.liter <- lm(logPrice ~ Liter, data = Cars_2005)
summary(logprice.liter)


```

### 2f.  Use linear regression to model log(Price) as a function of Liter + Leather.
```{r}

# with Leather
fitlogPrice = glm(logPrice ~ Liter + Leather, data = Cars_2005)
fitlogPrice

```

Write an equation for predicted log(Price) for cars without leather.

```{r}
# without Leather
fitlogPricewoLiter = glm(logPrice ~ Liter, data = Cars_2005)
fitlogPricewoLiter


```

Write an equation for predicted log(Price) for cars with leather.

### 2g.  Use logistic regression to model Cruise as a function of log(Price).
```{r}

fit = glm(Cruise ~ logPrice, data = Cars_2005, family = "binomial")
fit

```

Write an equation for predicted probability that a car will have cruise control.  Simplify your equation using the rules $e^{a+b} = e^a \cdot e^b$, $e^{ab}={(e^{a})}^b$ and $e^{\log(x)}=x$.

### 2h. Create a new data frame, cars2, containing the data from `cars`, plus a column containing the predicted probability that each car will have cruise control.
```{r}

```

### 2i.  Make a graph of the predicted probability of Cruise as a function of log(Price).
```{r}

```

Using the graph, estimate the probability that a car will have cruise control if its log(Price) is 9.5.  What if its log(Price) is 10.5?