---
title: "Lesson 7 Homework--DS 700"
author: "Abra Brisbin"
date: "12/18/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load necessary libraries here.
```{r}

library(ggformula)
library(dplyr)
library(tidyverse)
library(gapminder)

```
This assignment will help you practice using dplyr and ggformula to make some more complex types of graphs, and dig deeper into what a data set can tell us.

## Problem 1:  Relationship between life expectancy and GDP
### 1a.
Install and load the `gapminder` package.  We'll analyze the `gapminder` data set, which contains data on different countries in many different years.

Extract the data from the year 2007.  Make a scatterplot of lifeExp (life expectancy) as a function of gdpPercap (per capita GDP), for this subset of data.  (Hint:  You can pipe the subsetted data directly into the graph function.)
```{r}

country_df = gapminder
plot_df <- 
  country_df %>%
  filter(year == 2007)

gf_point(lifeExp ~ gdpPercap, data = plot_df) %>% 
  gf_lm() %>% 
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")

```

### 1b.
`gdpPercap` is right-skewed, so it may be helpful to use a log scale.  Pipe the results of your graph into `gf_refine(scale_x_log10())` to change the x-axis to a log scale.  

Use a log scale for gdpPercap throughout the rest of this assignment.
```{r}

plot_df <- plot_df %>%
  mutate(loggdpPercap=log(gdpPercap))

gf_point(lifeExp ~ loggdpPercap, data = plot_df) %>% 
  gf_lm() %>% 
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")

```

### 1c.
Modify your graph from 1b to use different colors for different continents.
```{r}
head(plot_df)

gf_point(lifeExp ~ loggdpPercap, col =~ continent, data = plot_df) %>%
  gf_lm() %>%
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")

```

### 1d.
Try adding a smoothing curve to the data for each continent.  What happens?  Why does this happen?
```{r}

gf_point(lifeExp ~ loggdpPercap, col =~ continent, data = plot_df) %>%
  gf_smooth() %>%
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")


```

### 1e.
One possible solution is to add a smoothing line, which is easier to compute than a loess curve.  Add a smoothing line to the data for each continent.
```{r}

gf_point(lifeExp ~ loggdpPercap, col =~ continent, data = plot_df) %>%
  gf_smooth(lifeExp ~ loggdpPercap, method = "line", col = "blue", data = plot_df) %>%
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")


```

### 1f.
An alternative solution is to use loess smoothing curves, but to focus on the continents with more data in 2007.  Extract the data from the year 2007 *and* from the four continents with the most data.  (The logical operators `&` (and) and `!=` (not equal to) may be helpful.)  Make a scatterplot of lifeExp versus gdpPercap, with smoothing curves for each continent.
```{r}

plottwo_df <- 
  country_df %>%
  filter(year == 2007 & continent != "Oceania")

head(plottwo_df)

gf_point(lifeExp ~ gdpPercap, data = plottwo_df) %>%
  gf_lm() %>%
  gf_labs(title = "How GDP of a country impacts Life Expectancy",
          x = "GDP per Capita",
          y = "Life Expectancy")

```

Add a title to your graph summarizing the key takeaway from the graph.  Add a caption stating the source of the data.
```{r}

```

## Problem 2:  Colors and labels
### 2a. 
Let's return to our graph from part 1c (without smoothing lines or curves).  We'd like to make our graph readable by people who are color-blind.  Visit https://colorbrewer2.org/ and select a color palette for 5 data classes that is "colorblind safe".  Then use the `scale_color_brewer()` function inside `gf_refine` to use the color palette you chose.  Experiment and find the palette you like the best!

Also map the shape of the points to the continent, to make the graph unambiguous for people who have printed your report in black and white.
```{r}

```

### 2b.
Use the function `coord_cartesian` (also inside `gf_refine`) to zoom in on the region of the graph with $10000 \leq GDP \leq 50000$ and $70 \leq lifeExp \leq 85$.
```{r}

gf_point(lifeExp ~ gdpPercap, col =~ continent, data = plot_df) %>%
  gf_smooth() + coord_cartesian(xlim = c(10000, 50000), ylim = c(70, 85)) 

summarycontinents <-
plot_df %>%
  filter(between(gdpPercap, 10000, 50000)  & between(lifeExp,70,85)) %>%
  group_by(continent) %>%
  summarise(numObs = n_distinct(country), LifeExpectancyMean = mean(lifeExp), GDPPerCapitaMean = mean(gdpPercap), PopulatioMedian = median(pop))
                
summarycontinents

```

### 2c.
Use `dplyr` to compute the mean life expectancy and mean GDP per capita of each continent in 2007.  Store your results in a data frame.
```{r}

plot_df <- 
  country_df %>%
  filter(year == 2007)

summarycontinents <-
plot_df %>%
  group_by(continent) %>%
  summarise(numObs = n_distinct(country), LifeExpectancyMean = mean(lifeExp), GDPPerCapitaMean = mean(gdpPercap), PopulatioMedian = median(pop))
                

summarycontinents

```

What was the mean life expectancy of African countries in 2007?  What was the mean GDP in Asian countries in 2007?

### 2d.
Use the data frame from part 2c, along with the function `gf_label`, to add continent labels to your graph from part 2a.  Make the labels partly transparent, so you can still see the data points underneath them.

- I recommend making the labels black, rather than using the ColorBrewer palette you used for the points, because the ligher colors can be hard to read.
```{r}

```

### 2e.
Return to the default color scheme.  Add the argument `fill =~ continent` to the labels to give the labels a background color that matches the points of the scatterplot.  Use the argument `show.legend = FALSE` in both the scatterplot and the labels to plot the graph without a legend.
```{r}

```
## Problem 3:  Population by continent
### 3a.
Extract the data from the year 2007 *and* from the four continents with the most data.  Make boxplots of the population of countries in each continent.  Use a log scale for population. 
```{r}

```

### 3b.
Use `dplyr` to compute the median population of countries in each continent in 2007.  Store your results in a data frame.
```{r}

```
Put the data frame in order by (increasing) median population.  Which continent is first in order?  Which continent is last?
```{r}

```

### 3c.
Re-do your graph from part 3a, but this time, put the boxplots in order of increasing median.
```{r}

```

### 3d.
Working with your graph from part 3c, increase the font size of the axis labels and titles.  You can do this using
`gf_theme(axis.text = element_text(...), axis.title = element_text(...))`

Read the documentation for the `element_text` function to determine the appropriate argument.
```{r}

```


